---
title: "Demo literature submission"
author: "Anthony Davidson"
date: "21/09/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
here::dr_here()
```

# Swift parrot ecology

This is a great species for this work as it includes challenging aspects of both observations and processes and how these play out in model space. 

## Background

There are several previous publications that test the following associations...

```{r}
#read_lit_db
```

## Conceptial model




## Challenges

Continous vs discrete data

```{r}



```


Precision vs. Accuracy



Abundance vs. Occupancy




Replication and psuedo-replication


There are the replicates? Well it depends on the scale you are interested in......

Because of the following aspects of statistics and mathematical projection.


## Methods

```{r}
#data
# source("./data/hollow_data_wrangled.xlsx", sheet = 1)
# simulate
# Generates 50 points from a N(0, 1): predictor and error
set.seed(34567)
x <- rnorm(n = 50)
eps <- rnorm(n = 50)

# Responses
yLin <- -0.5 + 1.5 * x + eps
yQua <- -0.5 + 1.5 * x^2 + eps
yExp <- -0.5 + 1.5 * 2^x + eps

# Data
leastSquares <- data.frame(x = x, yLin = yLin, yQua = yQua, yExp = yExp)

```


## Data and analysis

Trees dataset: 50m from nest trees

Hollow data = 1x nest hole +
              1x rand hole
              
Treelevel data

- nest, health

```{r}

# paperscriptcopy
#AnthonyDavidson
#21092021

library(tidyverse)
library(ggplot2)
library(effects)
library(lsmeans)
library(scales)
library(ggiraphExtra)
library(devtools)
library(car)
library(ggeffects)
library(readr)
library(readxl)

hollows<-read_excel("data/hollow_data_wrangled.xlsx")
trees<-read_excel("data/tree_data_wrangled.xlsx")

trees$health<-factor(trees$healthScale)

trees$healthScale<-factor(trees$healthScale)
#summaries
str(trees)
summary(trees)


str(hollows)
summary(hollows)

#nested trees 
nest <- factor(hollows$BirdPresence)
dbh <- as.numeric(hollows$treeDiameter)

# mod <- glm(nest~dbh)
# anova(summary(mod))



```

## Error modelling ... chancing variance

Catagorical vs continuous

### Import and wrangle data

```{r results = "asis"}
source("R/import_and_wrangle_data.R")

```

### Method

```{r}

```

### Results


```{r}
####QUESTION 1: IS THERE A DIFFERENCE IN THE NUMBER OF ENTRANCES DETECTED FROM THE GROUND VS. CLIMBING?####

#model choices....



#model selection

# errorn<-glm(BirdPresence~1,family = poisson(), data = trees)
# error1<-glm(groundcount~climbentry*DBH,family = poisson(), data = trees)#interaction is signfificant
# error2<-glm(groundcount~climbentry+DBH,family = poisson(), data = trees)
# error3<-glm(groundcount~climbentry,family = poisson(), data = trees)
# error4<-glm(groundcount~DBH,family = poisson(), data = trees)
# AIC(errorn,error1,error2,error3,error4)
# ##        df      AIC
# ## errorn  1 422.0013
# ## error1  4 319.0683
# ## error2  3 322.4391
# ## error3  2 324.9321
# ## error4  2 404.3212
# 322.4391-319.0683
# ## [1] 3.3708
# #we select model 'error1'
# summary(error1) #all terms significant

```


