---
title: "Gelfond's constant"
author: "Anthony Davidson and many others"
date: "27/09/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Gelfond's constant

```{r pressure, echo=FALSE}

# function(){
#  exp(pi)
# plot(pressure) 
# }

```

Gelfond's constant was named after Aleksandr Gelfond, Like both $\mvar \given | \e$ and $\pi$, this constant is a transcendental number. This was first established by Gelfond and may now be considered as an application of the [[Gelfond–Schneider theorem]], noting that ...

<!-- is {{math|''e''{{sup|''π''}}}}, that is, {{mvar|[[e (mathematical constant)|e]]}} raised to the [[exponentiation|power]]  [[Pi|{{pi}}]]. Like both {{mvar|e}} and {{pi}}, this constant is a [[transcendental number]]. This was first established by Gelfond and may now be considered as an application of the [[Gelfond–Schneider theorem]], noting that -->

$$e^\pi = (e^{i\pi})^{-i} = (-1)^{-i},$$

<ref>{{cite journal|author=Nesterenko, Y|author-link=Yuri Valentinovich Nesterenko|title=Modular Functions and Transcendence Problems|journal=[[Comptes rendus de l'Académie des sciences|Comptes Rendus de l'Académie des Sciences, Série I]]|volume=322|pages=909–914|year=1996|issue=10| zbl=0859.11047 }}</ref>

<!-- where {{mvar|i}} is the [[imaginary unit]]. Since {{math|&minus;''i''}} is algebraic but not rational, {{math|''e''{{sup|''π''}}}} is transcendental. The constant was mentioned in [[Hilbert's seventh problem]].<ref name=tijdeman>{{cite book | editor=Felix E. Browder | editor-link=Felix Browder | title=Mathematical Developments Arising from Hilbert Problems | series=[[Proceedings of Symposia in Pure Mathematics]] | volume=XXVIII.1 | year=1976 | publisher=[[American Mathematical Society]] | isbn=0-8218-1428-1 | first=Robert | last=Tijdeman | author-link=Robert Tijdeman | chapter=On the Gel'fond–Baker method and its applications | pages=241–268 | zbl=0341.10026 }}</ref> A related constant is {{math|2{{sup|{{sqrt|2}}}}}}, known as the [[Gelfond–Schneider constant]]. The related value {{pi}}&nbsp;+&nbsp;{{math|''e''{{sup|''π''}}}} is also irrational.< -->

<!-- ==Numerical value== -->
<!-- The decimal expansion of Gelfond's constant begins -->

<!-- :<math>e^\pi = </math> {{val|23.1406926327792690057290863679485473802661062426002119934450464095243423506904527835169719970675492196}}... &nbsp; {{OEIS2C|id=A039661}} -->

<!-- === Construction === -->
<!-- If one defines {{math|''k''{{sub|0}} {{=}} {{sfrac|1|{{sqrt|2}}}}}} and -->

<!-- :<math>k_{n+1} = \frac{1 - \sqrt{1 - k_n^2}}{1 + \sqrt{1 - k_n^2}}</math> -->

<!-- for {{math|''n'' > 0}}, then the sequence<ref>{{cite book|author1-link=Jonathan Borwein | last1=Borwein | first1=J. | last2=Bailey | first2=D. | title=Mathematics by Experiment: Plausible Reasoning in the 21st Century|url=https://archive.org/details/mathematicsbyexp00borw |url-access=limited |place=Wellesley, MA|publisher=A K Peters|year=2004|page=[https://archive.org/details/mathematicsbyexp00borw/page/n150 137] | isbn=1-56881-211-6 | zbl=1083.00001 }}</ref> -->

<!-- :<math>(4/k_{n+1})^{2^{-n}}</math> -->

<!-- converges rapidly to {{math|''e''{{sup|''π''}}}}. -->

<!-- == Continued fraction expansion == -->
<!-- :<math>e^{\pi} = 23+ -->
<!-- \cfrac{1} -->
<!--    {7+\cfrac{1} -->
<!--       {9 +\cfrac{1} -->
<!--          {3+\cfrac{1} -->
<!--             {1+\cfrac{1} -->
<!--                {1 +\cfrac{1} -->
<!--             {591+\cfrac{1} -->
<!--             {2+\cfrac{1} -->
<!--             {9+\cfrac{1} -->
<!--             {1+\cfrac{1} -->
<!--             {2+\cfrac{1} -->
<!--                {\ddots} -->
<!--                } -->

<!--             } -->

<!--             } -->

<!--             } -->
<!--                   } -->
<!--                } -->
<!--             } -->
<!--          } -->
<!--       } -->
<!--    } -->
<!-- </math> -->

<!-- This is based on the digits for the [[Continued fraction|simple continued fraction]]: -->

<!-- :<math>e^{\pi} = [23; 7, 9, 3, 1, 1, 591, 2, 9, 1, 2, 34, 1, 16, 1, 30, 1, 1, 4, 1, 2, 108, 2, 2, 1, 3, 1, 7, 1, 2, 2, 2, 1, 2, 3, 2, 166, 1, 2, 1, 4, 8, 10, 1, 1, 7, 1, 2, 3, 566, 1, 2, 3, 3, 1, 20, 1, 2, 19, 1, 3, 2, 1, 2, 13, 2, 2, 11, ...] -->
<!-- </math> -->

<!-- As given by the integer sequence [[oeis:A058287|A058287]]. -->

<!-- ==Geometric property== -->
<!-- The [[Volume of an n-ball|volume of the ''n''-dimensional ball]] (or [[n-ball|''n''-ball]]), is given by -->

<!-- :<math>V_n = \frac{\pi^\frac{n}{2}R^n}{\Gamma\left(\frac{n}{2} + 1\right)},</math> -->

<!-- where {{mvar|R}} is its radius, and {{math|Γ}} is the [[gamma function]]. Any even-dimensional ball has volume -->

<!-- :<math>V_{2n} = \frac{\pi^n}{n!}R^{2n},</math> -->

<!-- and, summing up all the unit-ball ({{math|''R'' {{=}} 1}}) volumes of even-dimension gives<ref>Connolly, Francis. University of Notre Dame{{full citation needed|date=November 2012}}</ref> -->

<!-- :<math>\sum_{n=0}^\infty V_{2n} (R = 1) = e^\pi.</math> -->

<!-- == Similar or related constants == -->

<!-- === Ramanujan's constant === -->
<!-- :<math>e^{\pi{\sqrt{163}}} = (\text{Gelfond's constant})^{\sqrt{163}}</math> -->

<!-- This is known as Ramanujan's constant. It is an application of [[Heegner number]]s, where 163 is the Heegner number in question. -->

<!-- Similar to {{math|''e''{{sup|''π''}} - ''π''}}, {{math|''e''{{sup|''π''{{sqrt|163}}}}}} is very close to an integer: -->

<!-- :<math>e^{\pi \sqrt{163}} = </math> {{val|262537412640768743.9999999999992500725971981856888793538563373369908627075374103782106479101186073129}}... <math> \approx 640\,320^3+744</math> -->

<!-- As it was the Indian mathematician [[Srinivasa Ramanujan]] who first predicted this almost-integer number, it has been named after him, though the number was first discovered by the French mathematician [[Charles Hermite]] in 1859. -->

<!-- The coincidental closeness, to within 0.000 000 000 000 75 of the number {{math|640320{{sup|3}} + 744}} is explained by [[complex multiplication]] and the [[Q-expansion|''q''-expansion]] of the [[j-invariant]], specifically: -->

<!-- :<math>j((1+\sqrt{-163})/2)=(-640\,320)^3</math> -->

<!-- and, -->

<!-- :<math>(-640\,320)^3=-e^{\pi \sqrt{163}}+744+O\left(e^{-\pi \sqrt{163}}\right)</math> -->

<!-- where {{math|''O''(''e''{{sup|-''π''{{sqrt|163}}}})}} is the error term, -->

<!-- :<math>{\displaystyle O\left(e^{-\pi {\sqrt {163}}}\right) = -196\,884/e^{\pi {\sqrt {163}}}\approx -196\,884/(640\,320^{3}+744)\approx -0.000\,000\,000\,000\,75}</math> -->

<!-- which explains why {{math|''e''{{sup|''π''{{sqrt|163}}}}}} is 0.000 000 000 000 75 below {{math|640320{{sup|3}} + 744}}. -->

<!-- (For more detail on this proof, consult the article on [[Heegner number]]s.) -->

<!-- === The number {{math|''e''{{sup|''π''}} - ''π''}} === -->
<!-- The decimal expansion of {{math|''e''{{sup|''π''}} - ''π''}} is given by [[oeis:A018938|A018938]]: -->

<!-- :<math>e^{\pi} - \pi = </math> {{val|19.9990999791894757672664429846690444960689368432251061724701018172165259444042437848889371717254321516}}... -->

<!-- Despite this being nearly the integer 20, no explanation has been given for this fact and it is believed to be a [[mathematical coincidence]]. -->

<!-- === The number {{math|''π''{{sup|''e''}}}} === -->
<!-- The decimal expansion of {{math|''π''{{sup|''e''}}}} is given by [[oeis:A059850|A059850]]: -->

<!-- :<math> -->
<!-- \pi^{e} = </math> {{val|22.4591577183610454734271522045437350275893151339966922492030025540669260403991179123185197527271430315}}... -->

<!-- It is not known whether or not this number is transcendental. Note that, by [[Gelfond–Schneider theorem|Gelfond-Schneider theorem]], we can only infer definitively that {{math|''a''{{sup|''b''}}}} is transcendental if {{mvar|a}} is algebraic and {{mvar|b}} is not rational ({{mvar|a}} and {{mvar|b}} are both considered [[complex number]]s, also {{math|''a'' ≠ 0}}, {{math|''a'' ≠ 1}}). -->

<!-- In the case of {{math|''e''{{sup|''π''}}}}, we are only able to prove this number transcendental due to properties of complex exponential forms, where {{pi}} is considered the modulus of the complex number {{math|''e''{{sup|''π''}}}}, and the above equivalency given to transform it into {{math|(-1){{sup|-''i''}}}}, allowing the application of Gelfond-Schneider theorem. -->

<!-- {{math|''π''{{sup|''e''}}}} has no such equivalence, and hence, as both {{pi}} and {{mvar|e}} are transcendental, we can make no conclusion about the transcendence of {{math|''π''{{sup|''e''}}}}. -->

<!-- === The number {{math|''e''{{sup|''π''}} - ''π''{{sup|''e''}}}} === -->
<!-- As with {{math|''π''{{sup|''e''}}}}, it is not known whether {{math|''e''{{sup|''π''}} - ''π''{{sup|''e''}}}} is transcendental. Further, no proof exists to show whether or not it is irrational. -->

<!-- The decimal expansion for {{math|''e''{{sup|''π''}} - ''π''{{sup|''e''}}}} is given by [[oeis:A063504|A063504]]: -->

<!-- :<math> -->
<!-- e^{\pi} - \pi^{e} = </math> {{val|0.6815349144182235323019341634048123526767911086035197442420438554574163102913348711984522443404061881}}... -->

<!-- === The number {{math|''i''{{sup|''i''}}}} === -->
<!-- :<math>i^{i} = (e^{i\pi/2})^i = e^{-\pi/2} = (e^{\pi})^{-1/2}</math> -->

<!-- The decimal expansion of is given by [[oeis:A049006|A049006]]: -->

<!-- :<math> -->
<!-- i^{i} = </math> {{val|0.2078795763507619085469556198349787700338778416317696080751358830554198772854821397886002778654260353}}... -->

<!-- Because of the equivalence, we can use Gelfond-Schneider theorem to prove that the reciprocal square root of Gelfond's constant is also transcendental: -->

<!-- {{mvar|i}} is both algebraic (a solution to the polynomial {{math|''x''{{sup|2}} + 1 {{=}} 0}}), and not rational, hence {{math|''i''{{sup|''i''}}}} is transcendental. -->

<!-- == See also == -->
<!-- * [[Transcendental number]] -->
<!-- * [[Transcendental number theory]], the study of questions related to transcendental numbers -->
<!-- * [[Euler's identity]] -->
<!-- * [[Gelfond–Schneider constant]] -->

<!-- ==References== -->
<!-- {{Reflist}} -->

<!-- ==Further reading== -->
<!-- * [[Alan Baker (mathematician)|Alan Baker]] and [[Gisbert Wüstholz]], ''Logarithmic Forms and Diophantine Geometry'', New Mathematical Monographs '''9''', Cambridge University Press, 2007, {{ISBN|978-0-521-88268-2}} -->

<!-- ==External links== -->
<!-- * [http://mathworld.wolfram.com/GelfondsConstant.html Gelfond's constant at ''MathWorld''] -->
<!-- * [https://web.archive.org/web/20091027100754/http://www.geocities.com/timeparadox/Muntekim.htm  A new complex power tower identity for Gelfond's constant] -->
<!-- * [http://mathworld.wolfram.com/AlmostInteger.html Almost Integer at ''MathWorld''] -->

<!-- [[Category:Mathematical constants]] -->
<!-- [[Category:Real transcendental numbers]] -->
